{\tt APFELgrid}: a fast interface to hadronic observables for fits of parton densities

### 2016-04-29 paper.tex
updating layout

@@

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

\title{{\tt APFELgrid}: a fast interface to hadronic observables for
  fits of parton densities}

@@

\cortext[author] {Corresponding author.\\\textit{E-mail address:} nathan.hartland@physics.ox.ac.uk}
\address[a]{Rudolf Peierls Centre for Theoretical Physics,\\ 1 Keble Road, University of Oxford, OX1 3NP, Oxford, UK}
\address[b]{TH Unit, CERN, CH-1211 Geneva 23, Switzerland}

@@

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
QCD; parton distribution functions; fast predictions.

\end{keyword}

@@

@@

%%\tableofcontents

@@

\begin{equation} \label{eq:applconv}
  \sigma_{pp\to X} = \sum_p^{N_\alpha} \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta,\tau} 
  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,.
\end{equation}
where
\begin{equation}
  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)\,,
\end{equation}
consists of the convolution of the hard cross-section with the interpolating polynomials. This information may be stored in a precomputed look-up table.
The final expression for the cross-section in Eqn.~\ref{eq:applconv} therefore being considerably more computationally efficient inside a fit than the direct
performance of the double-convolution.

@@

The {\tt APFELgrid} plugin generates {\tt FK} tables of the form expressed in Eqn.~\ref{eq:FKTable} from weight grids supplied by {\tt APPLgrid} files. In this section we shall examine the performance differences between the two expressions for cross-section prediction Eqn.~\ref{eq:applconv} ({\tt APPLgrid}) and Eqn.~\ref{eq:FK} ({\tt FK}). For a comprehensive benchmark, we shall consider a wide range of processes including LHC and Tevatron electroweak vector boson production measurements, inclusive jet data and $t\bar{t}$ total cross-sections. Predictions are provided for a wide range of kinematics, for a total of 52 source {\tt APPLgrid} files roughly corresponding to the full collider dataset of the forthcoming NNPDF3.1 PDF determination.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ph]
  \centering
  \hspace*{-2.0cm}                                                           
  \includegraphics[scale=0.5]{plots/t0}
\caption{\small Performance comparisons between {\tt APFELgrid} with
  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
  convolution time per point and process.}
\label{fig:benchmark}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ph]
  \centering
  \includegraphics[scale=0.5]{plots/t0a}
\caption{\small Performance comparisons between {\tt APFELgrid} with
  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
  convolution time per point and process.}
\label{fig:benchmark}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ph]
  \centering                                      
  \includegraphics[scale=0.5]{plots/t0b}
\caption{\small Performance comparisons between {\tt APFELgrid} with
  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
  convolution time per point and process.}
\label{fig:benchmark}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\textwidth=16.0cm \textheight=21.0cm 
\topmargin 0cm \oddsidemargin 0cm 
\setlength{\unitlength}{1mm}

@@

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{url,hyperref}

%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

\title{{\tt APFELgrid}: a fast interface to hadronic observables for
  fits of parton densities}
\tnotetext[mytitlenote]{Preprints: OUTP-16-09P, CERN-TH-2016-103}

@@

\cortext[author] {Corresponding author.\\\textit{E-mail address:} nathan.hartland@physics.ox.ac.uk}
\address[a]{Rudolf Peierls Centre for Theoretical Physics,\\ 1 Keble Road, University of Oxford, OX1 3NP, Oxford, UK}
\address[b]{Theory Department, CERN, CH-1211 Geneva 23, Switzerland}

@@

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
QCD; parton distribution functions; fast predictions.
\end{keyword}

@@

 \newpage

\noindent {\Large \textbf{Program Summary}}\\

@@

\tableofcontents

@@

\begin{equation} \label{eq:applconv}
  \sigma_{pp\to X} = \sum_p^{N_\alpha} \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta,\tau} 
  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,.
\end{equation}
where
\begin{equation}
  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)\,,
\end{equation}
consists of the convolution of the hard cross-section with the interpolating polynomials. This information may be stored in a precomputed look-up table.
The final expression for the cross-section in Eq.~(\ref{eq:applconv}) therefore being considerably more computationally efficient inside a fit than the direct
performance of the double-convolution.

@@

The {\tt APFELgrid} plugin generates {\tt FK} tables of the form expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt APPLgrid} files. In this section we shall examine the performance differences between the two expressions for cross-section prediction Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt FK}). For a comprehensive benchmark, we shall consider a wide range of processes including LHC and Tevatron electroweak vector boson production measurements, inclusive jet data and $t\bar{t}$ total cross-sections. Predictions are provided for a wide range of kinematics, for a total of 52 source {\tt APPLgrid} files roughly corresponding to the full collider dataset of the forthcoming NNPDF3.1 PDF determination.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[p]
  \centering
  \includegraphics[scale=0.5]{plots/t0a}
\caption{\small Performance comparisons between {\tt APFELgrid} with
  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
  convolution time per point and process.}
\label{fig:benchmark}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[p]
  \centering                                      
  \includegraphics[scale=0.5]{plots/t0b}
\caption{\small Performance comparisons between {\tt APFELgrid} with
  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
  convolution time per point and process.}
\label{fig:benchmark}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section*{Acknowledgements}

We thanks the NNPDF Collaboration members for the support and
motivation they gave us to write this work. V.~B. and N.~H. are
supported by an European Research Council Starting Grant ``PDF4BSM''.
S.~C. is supported by the HICCUP ERC Consolidator grant (614577).
